---
title: "Colony Forming Units assay"
author: "Liz Hughes & Edward Wallace"
date: "31/10/2022"
output:
  html_document:
    toc: yes
    toc_depth: 2
---

```{r call_shared_functions_formatting, child = "shared_functions_formatting.Rmd"}
```

# Summary

Collated OD measurements and average cell counts (from CFU assays).

2022-10-31 Comparison of growth of H99 and gat201∆m mutant (Madhani collection) and 2 Gat201 complemented strains (GAT201-C1, formerly C26; GAT201-C2, formerly C30) into the Madhani deletion background in RPMI and RPMI + Serum (10% FBS) at 37°C over a 60 hour time course.

Samples were collected at 4, 12, 24, 36, 48 and 60 hours post inoculation into fresh media. A 1:10 dilution series was carried out from neat to 10e-4. 100 microlitres of dilutions 10e3 and 10e4 were plated out onto YPD Agar, incubated for 48 hours and colonies counted.

The average of cells/ml over the 2 dilutions were used as input for the data analysis.


## Read in data as csv file

```{r nice_names}
strain_df <- tibble( Strain      = c("WT", "Gat201 del", "C26", "C30" ),
                     Strain_nice = c("GAT201", "gat201∆m", "GAT201-C1", "GAT201-C2"),
                     colour      = c("grey20", "darkred", "blue3", "cyan4"))

scale_color_strain <- 
  scale_color_manual("Strain",
                     breaks = strain_df$Strain_nice,
                     values = strain_df$colour)
```

```{r load_Cfu_data}

cfu_data <- 
  here::here("data", "20221025_CFU60_RPMI_RPMIserum.csv") %>%
  readr::read_csv(comment = "#") %>%
  left_join(strain_df, by = "Strain") %>%
  mutate(Time_days   = Time/24, 
         BioRep      = as_factor(BioRep),
         Strain      = as_factor(Strain),
         Strain_nice = as_factor(Strain_nice))
```



## Create Cfu plot
### First look at data.


```{r plot_all}
Cfu_Plot <- 
  ggplot(cfu_data,
         aes(x = Time_days, y = Cells, colour = Media)) +
  geom_point(aes(shape = BioRep), size=3) +
  labs(x="Time (Days)", 
       y="Number of Cells/ml", 
       title = "CFU assay at 37°C") +
  scale_x_continuous(expand = c(0, 0), breaks = 0:3, limits = c(0,3)) + 
  scale_y_continuous(breaks = c(10, 100000,  1000000, 10000000)) +
  facet_wrap(~Strain_nice)

Cfu_Plot
```



## Plot mean CFU


```{r plot_all_mean_dash, fig.height = 4, fig.width = 3.75}

Cfu_Plot2 <- 
  ggplot(cfu_data,
         aes(x = Time, y = Cells, colour = Strain_nice)) +
  geom_point(aes(shape = BioRep), size=1) +
  scale_color_strain +
  facet_wrap(~ Media, ncol = 1, scales = "free_x") +
  labs(x = "Time (hours)", 
       y = "Number of Colonies/ml",
       shape = "Replicate") +
  scale_x_continuous(expand = c(0, 0),
                     breaks = c(0, 4, 12, 24, 36, 48, 60)) + 
  scale_y_continuous(trans = scales::pseudo_log_trans(sigma =100000,
                                                    base =  exp(1)),
                     breaks = c(0, 100000, 500000, 1000000, 5000000),
                     expand = c(0, 0),
                    labels = parse(text = c("0", "10^5", "5 %*% 10^5",  "10^6", "5 %*% 10^6"))) +
  coord_cartesian(clip = "off") +
  guides(colour = guide_legend(order = 1),
         shape = guide_legend(order = 2)) +
  theme(panel.border = element_blank())

Cfu_Plot2_mean <-  Cfu_Plot2 + 
  stat_summary(aes(group = Strain), 
               fun = mean, geom="line", 
               linetype = "longdash", size = 0.6)

Cfu_Plot2_mean
```


```{r mean_plot_all_RRS.png}
ggsave("../Results/cfu_mean_plot_all_RRS.png",
       width = 3.75, height = 4)
```

```{r mean_plot_all_RRS.svg}
ggsave("../Results/cfu_mean_plot_all_RRS.svg",
       width = 3.75, height = 4)
```


# Display session info with package versions

To aid reproducibility and troubleshooting.

```{r sessionInfo}
sessionInfo()
```

