---
title: "Plate reader growth curves of GAT204, LIV3 and double deletion mutants"
author: "Liz Hughes & Edward Wallace"
date: "05/04/2024"
output:
  html_document:
    toc: yes
    toc_depth: 2
---

```{r call_shared_functions_formatting, child = "shared_functions_formatting.Rmd"}
```

# Summary

This experiments tests the effect of gat204∆, liv3∆, and the double deletion, on C. neoformans growth in RPMI media. It uses wild-type strain and gat201∆ as controls.

This concludes that (again) both gat204∆ and liv3∆ alone have intermediate growth phenotype between wild-type and gat201∆ cells.
New here, the double deletion mutant gat204∆liv3∆ has an intermediate phenotype similar to both those single mutants.

Data in files, one per biological replicate plate:

- `20240403_GAT204LIV3.xlsx`


# Set up data frame of nice strain names, colours, styles.


```{r strain_df}
strain_df <- tibble( Strain      = c("KN99-1", "Gat201(M)", "Gat204", "Liv3", "Gat201-Liv3"),
                     Strain_nice = c("WT", "gat201∆", "gat204∆", "liv3∆", "gat204∆liv3∆"),
                     colour      = c("grey20", "darkred", "#d95f02", "#1b9e77",   "purple"))
```


## Read in the Plate map data from csv file RPMI

Also combine with nice strain names for display.

```{r load_platemap}
platemap <- 
  here::here("data", "20240403_GAT204LIV3_Setup.csv") %>%
  read_platemap_csv(strain_df = strain_df)

head(platemap, n=10)

```

# Analysis, biological replicate 1

## Read in raw data

```{r load_raw_od}
raw_od <-
  here::here("data", "20240403_GAT204LIV3.xlsx") %>%
  read_platereader_xlsx_magellan(well_names = well_names_96_colfirst)

```



## Reshape data and combine with the plate map, pairing them by Well


```{r reshape_annotate}
annotated_od <- 
  reshape_annotate_raw_od(raw_od = raw_od, platemap = platemap) %>%
  mutate(Well = factor(Well, levels = well_names_96_rowfirst))

```

## Plot ODs from all wells

```{r plot_allwells, fig.height = 6, fig.width = 12}
ggplot(data=annotated_od,
       aes(x=Time, y=OD595,  group = Well, colour = Strain)) + 
  geom_line() + 
  scale_time_hrs(limits = c(0,24)) +
  scale_color_strain(strain_df) + 
  facet_wrap(~ Well, ncol = 12) 
```

## Plot blank ODs

To find how stable the ODs of the blank wells are.

```{r plot_blank_ODs, fig.width = 5.5, fig.height = 4}
ggplot(data=filter(annotated_od, is.na(Strain)),
       aes(x=Time, y=OD595,  group = Well, colour = Medium)) + 
  geom_line() + 
  scale_time_hrs(limits = c(0,24)) + 
  scale_y_continuous(limits = c(0, 0.25),
                     expand = c(0, 0),
                     breaks = c(0, 0.05, 0.1, 0.15, 0.2)) +
  labs(title = "Blank wells", x="Time (hours)", y="raw OD595")
```

This shows that OD of blank wells is somewhat messy, but within reasonable bounds.


## Calculate median OD for blank wells


```{r calc_blank_medians}
blank_od_summary <- 
  annotated_od %>%
  dplyr::group_by(Medium) %>%
  summarise_od_fixed(od.name = "OD595",
                     is.na(Strain)) 

print(blank_od_summary)

blank_od_time <- 
  annotated_od %>%
  dplyr::group_by(Medium) %>%
  summarise_od_bytime(od.name = "OD595",
                     is.na(Strain),
                     !is.na(OD595)) 

print(blank_od_time)
```

## Subtract blank OD to make corrected OD

```{r subtract_blank_OD}
normalised_od <- 
  annotated_od %>%
  normalise_od_2ways(is.na(Strain))
```

## Plot OD corrected


```{r plot_all_stat_summary, fig.width = 7, fig.height = 5}
plot <- 
  ggplot_growthcurve_summary(data_od = normalised_od %>%
                               filter(!is.na(Strain)),
                             value_od = "OD_corrected_time",
                             limits_time = c(0,24),
                             limits_od = NULL,
                             breaks_od = c(0, 0.05, 0.1, 0.2, 0.3, 0.4, 0.5)) +
  scale_color_strain(strain_df) +
  expand_limits(y = c(0, 0.12))

plot
```

Conclude that the phenotype of gat201∆liv3∆ double mutant is no more severe than that of either single mutant.


## Plot OD corrected, with smoothing spline


```{r plot_all_stat_spline, fig.width = 7, fig.height = 5}
spline_plot <- 
  ggplot_growthcurve_spline(data_od = normalised_od %>%
                               filter(!is.na(Strain)),
                             value_od = "OD_corrected_time",
                             limits_time = c(0,24),
                             limits_od = NULL,
                             breaks_od = c(0, 0.05, 0.1, 0.15, 0.2)) +
  scale_color_strain(strain_df) +
  expand_limits(y = c(0, 0.2))

spline_plot
```

```{r save_plot_allreps}
ggsave(filename = here::here("results", "fig_growthcurves_GAT204_LIV3.png"), 
       plot = plot, height = 5, width = 6)
```



# Session info for reproducibility

```{r session_info}
sessionInfo()
```



















