---
title: "Plate reader growth curves of GAT201 mutants in RPMI media combining NaHCO3 and cAMP"
author: "Liz Hughes & Edward Wallace"
date: "05/03/2024"
output:
  html_document:
    toc: yes
    toc_depth: 2
---

```{r call_shared_functions_formatting, child = "shared_functions_formatting.Rmd"}
```

# Summary

This experiments tests the effect of sodium bicarbonate (NaHCO3) and cAMP on growth of Cryptococcus neoformans.

Plate reader growth curves of  WT KN99-alpha and gat201∆ deletion mutant Madhani in RPMI media with no phenol red.

- 2 Bioreps (1 per plate), 3 Techreps each. One condition has has 1 Biorep with 3 Techreps
- Grow 5ml culture from colony (1 colony = 1 Biorep) in YPD, 30C, 180 rpm ON.
- Seed at OD 600nm = 0.2 , 200 ul per well.
- Run for 1 day / 24 hours at 37C
- With 2 concentrations of NaHCO3: 0mM, 24mM
- With 2 concentrations of dbcAMP: 0mM (water control), 10 mM; or 10mM sodium butyrate (Nabut) to control for butyrate toxicity as a breakdown product of dbcAMP.

Data in files, one per biological replicate plate:

- `20240305_cAMP_Assay_TRSP.csv`

# Set up data frame of nice strain names, colours, styles.


```{r strain_df}
strain_df <- tibble( Strain      = c("KN99-1", "KN99-2", "Gat201(M)", "Gat201(M)-2"),
                     Strain_nice = c("GAT201", "GAT201", "gat201∆m", "gat201∆m"),
                     colour      = c("grey20", "grey20", "darkred", "darkred"))
```

```{r media_df}
media_df <- tibble( Medium      = as_factor(c("RPMI-N", "RPMI-24", "RPMI-N+W", "RPMI-24+W", "RPMI-N+C", "RPMI-24+C", "RPMI-N+SB", "RPMI-24+SB" )),
                    NaHCO3      = as_factor(c("−", "+", "−", "+", "−", "+", "−", "+" )),
                    cAMP      = as_factor(c("−", "−", "−", "−", "+", "+", "SB", "SB" )),
                    addition      = as_factor(c("H2O", "H2O", "H2O", "H2O", "dbcAMP", "dbcAMP", "Nabut", "Nabut" ))
                    )
```

## Read in the Plate map data from csv file RPMI

Also combine with nice strain names for display.

```{r load_platemap}
platemap <- 
  here::here("data", "20240305_cAMP_Plate_Setup.csv") %>%
  read_platemap_csv(strain_df = strain_df) %>%
  left_join(media_df, by = "Medium")

head(platemap, n=10)

```

# Analysis, biological replicate 1

## Read in raw data

```{r load_raw_od}
raw_od <-
  here::here("data", "20240305_cAMP_Assay.xlsx") %>%
  read_platereader_xlsx_magellan()

# raw_od <-
#   here::here("data", "20240305_cAMP_Assay_TRSP.csv") %>%
#   read_platereader_csv()
```



## Reshape data and combine with the plate map, pairing them by Well


```{r reshape_annotate}
annotated_od <- 
  reshape_annotate_raw_od(raw_od = raw_od, platemap = platemap)  %>%
  mutate(Well = factor(Well, levels = well_names_96_rowfirst))

```

## Plot ODs from all wells

```{r plot_allwells, fig.height = 6, fig.width = 12}
ggplot(data=annotated_od,
       aes(x=Time, y=OD595,  group = Well, colour = Strain)) + 
  geom_line() + 
  scale_time_hrs(limits = c(0,24)) +
  scale_color_strain(strain_df) + 
  facet_wrap(~ Well, ncol = 12) 
```

## Plot blank ODs

To find how stable the ODs of the blank wells are.

```{r plot_blank_ODs, fig.width = 5.5, fig.height = 4}
ggplot(data=filter(annotated_od, is.na(Strain)),
       aes(x=Time, y=OD595,  group = Well, colour = Medium)) + 
  geom_line() + 
  scale_time_hrs(limits = c(0,24)) + 
  scale_y_continuous(limits = c(0, 0.25),
                     expand = c(0, 0),
                     breaks = c(0, 0.05, 0.1, 0.15, 0.2)) +
  labs(title = "Blank wells", x="Time (hours)", y="raw OD595")
```

This shows that OD of blank wells is somewhat messy, but within reasonable bounds.


## Calculate median OD for blank wells


```{r calc_blank_medians}
blank_od_summary <- 
  annotated_od %>%
  dplyr::group_by(Medium) %>%
  summarise_od_fixed(od.name = "OD595",
                     is.na(Strain)) 

print(blank_od_summary)

blank_od_time <- 
  annotated_od %>%
  dplyr::group_by(Medium) %>%
  summarise_od_bytime(od.name = "OD595",
                     is.na(Strain),
                     !is.na(OD595)) 

print(blank_od_time)
```

## Subtract blank OD to make corrected OD

```{r subtract_blank_OD}
normalised_od <- 
  annotated_od %>%
  normalise_od_2ways(is.na(Strain))
```

## Plot OD corrected


```{r plot_all_stat_summary, fig.width = 7, fig.height = 6}
plot <- 
  ggplot_growthcurve_summary(data_od = normalised_od %>%
                               filter(!is.na(Strain)),
                             value_od = "OD_corrected_fixed",
                             limits_time = c(0,24),
                             limits_od = NULL,
                             breaks_od = c(0, 0.05, 0.1, 0.2, 0.3, 0.4, 0.5)) +
  scale_color_strain(strain_df) +
  expand_limits(y = c(0, 0.12)) +
  facet_rep_grid(NaHCO3 ~ addition, 
                 scales = "free_y", 
                 space = "free_y",
                 labeller = label_both)

plot
```

## Same unsmoothed plot for supplementary figure

Conclude that dbcAMP does not qualitatively affect the phenotype of GAT201 affecting growth in a NaHCO3-dependent way.

Sodium butyrate as additional negative control also does not strongly affect growth.

## Plot OD corrected, with smoothing spline


```{r plot_all_stat_spline, fig.width = 7, fig.height = 6}
spline_plot <- 
  ggplot_growthcurve_spline(data_od = normalised_od %>%
                               filter(!is.na(Strain)),
                             value_od = "OD_corrected_fixed",
                             limits_time = c(0,24),
                             limits_od = NULL,
                             breaks_od = c(0, 0.05, 0.1, 0.2, 0.3, 0.4, 0.5)) +
  scale_color_strain(strain_df) +
  expand_limits(y = c(0, 0.12)) +
  facet_rep_grid(NaHCO3 ~ addition, 
                 scales = "free_y", 
                 space = "free_y",
                 labeller = label_both)

spline_plot
```

```{r save_plot_allreps}
ggsave(filename = here::here("results", "fig_growthcurves_NaHCO3_vs_cAMP.png"), 
       plot = plot, height = 7, width = 6)
```



# Session info for reproducibility

```{r session_info}
sessionInfo()
```



















