---
title: "Plate reader growth curves of GAT201 targets in RPMI"
author: "Liz Hughes & Edward Wallace"
date: "16/05/2023"
output:
  html_document:
    toc: yes
    toc_depth: 2
---

```{r call_shared_functions_formatting, child = "shared_functions_formatting.Rmd"}
```

# Summary

Plate reader growth curves of GAT201 targets, March 2023

# Plate 1, 20230308, for main panel figure

Design:

- WT KN99-alpha, gat201∆, gat204∆, liv3∆, blp1∆, all from Madhani deletion collection
- 4 Biological replicates, 3 Technical replicates each.
- Grow 5ml culture from colony (1 colony = 1 Biorep) in YPD, 30C, 180 rpm ON.
- Seed at OD 600nm = 0.2 , 200 ul per well.
- Grow in RPMI at 37 degrees C.
- Run for 2 days (283 cycles).


# Set up data frame of nice strain names, colours, styles.

```{r strain_df_1}
strain_df_1 <- tibble( 
                    Strain = c("KN99 (WT)", "GAT201", "GAT204", "LIV3", "BLP1"),
                    Strain_nice =  c("WT", "gat201∆", "gat204∆", "liv3∆", "blp1∆"),
                    colour = c("grey20", "darkred", "#d95f02", "#1b9e77",   "#7570b3"))
```


## Read in transposed data as csv file

```{r load_raw_od}
raw_od <-
  here::here("data", "20230308_RPMI_Gat201Targets_Plate1.xlsx") %>%
  read_platereader_xlsx(start_row = 52)
```

## Read in the Plate map data from csv file

Also combine with nice strain names for display.

```{r load_platemap}
platemap <- 
  here::here("data", "20230308_RPMI_Gat201Targets_Plate1_setup.csv") %>%
  read_platemap_csv(strain_df = strain_df_1) 

head(platemap, n=10)

```

## Reshape data and combine with the plate map, pairing them by Well

```{r reshape_annotate}
annotated_od <- 
  reshape_annotate_raw_od(raw_od = raw_od, platemap = platemap)
```

## Plot blank ODs

To find how stable the ODs of the blank wells are.

```{r plot_blank_ODs, fig.width = 5, fig.height = 4}
ggplot_growthcurve(data_od=filter(annotated_od, is.na(Strain)),
                   limits_time = c(0, 48),
                   limits_od = c(0, 0.2),
                   breaks_od = c(0, 0.05, 0.1, 0.15)) + 
  geom_line(aes(group = Well)) + 
  labs(title = "Blank wells RPMI")
```

This shows that the OD increases over time, presumably pH change of the phenol red indicator. We correct for this.

## Plot raw ODs of all wells

```{r plot_all_wells, fig.width = 5, fig.height = 4}
ggplot_growthcurve(data_od = annotated_od,
                   limits_time = c(0, 48),
                   limits_od = c(0, 0.3),
                   breaks_od = c(0, 0.05, 0.1, 0.15, 0.2, 0.25)) +
  geom_line(aes(group = Well, colour = Strain))
```

## Calculate median OD for blank wells

```{r calc_blank_medians}
# filter out columns 1 and 12 as well?
blank_od_summary <- 
  summarise_od_fixed(annotated_od, od.name = "OD595",
                     is.na(Strain)) 

print(blank_od_summary)

blank_od_time <- 
  summarise_od_bytime(annotated_od, od.name = "OD595",
                     is.na(Strain)) 

print(blank_od_time)
```

## Subtract blank OD to make corrected OD

```{r subtract_blank_OD}
normalised_od <- 
  normalise_od_2ways(annotated_od, is.na(Strain))
```

## Plot OD corrected by time-based median summary

```{r plot_all_stat_summary, fig.width = 5, fig.height = 3.75}
plot_plate_1 <- 
  ggplot_growthcurve_spline_pluswells(data_od = normalised_od %>%
                               filter(!is.na(Strain)),
                             value_od = "OD_corrected_time",
                             limits_time = c(0, 48),
                             limits_od = c(0, 0.11)) +
  scale_color_strain(strain_df_1)

plot_plate_1
```


```{r save_png_summary}
here::here("results",
           "20230308_RPMI_Gat201Targets_Plate1_summaryplot.png") %>%
  ggsave(plot = plot_plate_1,
         width = 5, height = 3.75)
```

## Plot OD corrected by time-based median summary, smoothed

```{r plot_all_stat_spline, fig.width = 5, fig.height = 3.75}
plot_plate_1_spline <- 
  ggplot_growthcurve_spline(data_od = normalised_od %>%
                              filter(!is.na(Strain)),
                            value_od = "OD_corrected_time",
                            limits_time = c(0, 48)) +
  scale_color_strain(strain_df_1) 

plot_plate_1_spline
```

```{r save_png_spline}
here::here("results",
           "20230308_RPMI_Gat201Targets_Plate1_splineplot.png") %>%
  ggsave(plot = plot_plate_1_spline,
         width = 5, height = 3.75)
```

```{r save_svg_spline}
here::here("results",
           "20230308_RPMI_Gat201Targets_Plate1_splineplot.svg") %>%
  ggsave(plot = plot_plate_1_spline + 
           theme(legend.position = "bottom") + 
           guides(colour = guide_legend(ncol = 1)),
         width = 3.3, height = 4)
```


# Plate 2: WT, gat201∆, ecm2201∆, mep1∆, pdr802∆

```{r load_plate_2, fig.width = 5, fig.height = 3.75}
strain_df_2 <- tibble( 
                    Strain = c("KN99 (WT)", "GAT201", "ECM2201", "MEP1", "PDR802"),
                    Strain_nice =  c("WT", "gat201∆", "ecm2201∆", "mep1∆", "pdr802∆"),
                    colour = c("grey20", "darkred", "#d95f02", "#1b9e77",   "#7570b3"))

platemap_2 <- 
  here::here("data", "20230310_RPMI_Gat201Targets_Plate2_setup.csv") %>%
  read_platemap_csv() 

annotated_od_2 <- 
  load_annotate_platereader(
    file_raw_od = "20230310_RPMI_Gat201Targets_Plate2.xlsx", 
    file_platemap = "20230310_RPMI_Gat201Targets_Plate2_setup.csv", 
    strain_df = strain_df_2)

ggplot_growthcurve(data_od = annotated_od_2,
                   limits_time = c(0, 48),
                   limits_od = c(0, 0.3),
                   breaks_od = c(0, 0.05, 0.1, 0.15, 0.2, 0.25)) +
  geom_line(aes(group = Well, colour = Strain))


normalised_od_2 <- annotated_od_2 %>%
  normalise_od_2ways(is.na(Strain))

```


## Plot OD corrected by time-based median summary

```{r plot_plate_2_stat_summary, fig.width = 5, fig.height = 3.75}
plot_plate_2 <- 
  ggplot_growthcurve_spline_pluswells(data_od = normalised_od_2 %>%
                               filter(!is.na(Strain)),
                             value_od = "OD_corrected_time",
                             limits_time = c(0, 48),
                             limits_od   = c(0, 0.11)) +
  scale_color_strain(strain_df_2)

plot_plate_2
```


# Plate 3: WT, gat201∆, CNAG_04874, hxt1∆, fcy2∆

Note, [CNAG_04874](https://fungidb.org/fungidb/app/record/gene/CNAG_04874) is homologous to [*S. pombe* Crr1](https://www.pombase.org/gene/SPBC21B10.07).

```{r load_plate_3, fig.width = 5, fig.height = 3.75}
strain_df_3 <- tibble( 
                    Strain = c("KN99 (WT)", "GAT201", "CNAG_04874", "HXT1", "FCY2"),
                    Strain_nice =  c("WT", "gat201∆", "cnag_04874∆", "hxt1∆", "fcy2∆"),
                    colour = c("grey20", "darkred", "#d95f02", "#1b9e77",   "#7570b3"))

platemap_3 <- 
  here::here("data", "20230321_RPMI_Gat201Targets_Plate3_setup.csv") %>%
  read_platemap_csv() 

annotated_od_3 <- 
  load_annotate_platereader(
    file_raw_od = "20230321_RPMI_Gat201Targets_Plate3.xlsx", 
    file_platemap = "20230321_RPMI_Gat201Targets_Plate3_setup.csv", 
    strain_df = strain_df_3)

ggplot_growthcurve(data_od = annotated_od_3,
                   limits_time = c(0, 48),
                   limits_od = c(0, 0.3),
                   breaks_od = c(0, 0.05, 0.1, 0.15, 0.2, 0.25)) +
  geom_line(aes(group = Well, colour = Strain))

normalised_od_3 <- annotated_od_3 %>%
  normalise_od_2ways(is.na(Strain))

```

## Plot OD corrected by time-based median summary

```{r plot_plate_3_stat_summary, fig.width = 5, fig.height = 3.75}
plot_plate_3 <- 
  ggplot_growthcurve_spline_pluswells(data_od = normalised_od_3 %>%
                               filter(!is.na(Strain)),
                             value_od = "OD_corrected_time",
                             limits_time = c(0, 48),
                             limits_od   = c(0, 0.11)) +
  scale_color_strain(strain_df_3)

plot_plate_3
```


# Summary figure Plates 1-3 (Supplementary Figure S11)

```{r plot_plates123, fig.width = 5, fig.height = 7.5}
plot_plates_123 <-
  plot_grid(plot_plate_1,
            plot_plate_2,
            plot_plate_3,
            align = "v",
            ncol = 1,
            labels = "AUTO")

plot_plates_123
```

```{r save_png_summary_plate123}
here::here("results",
           "fig_growthcurves_Gat201Targets_RPMI_summary.png") %>%
  ggsave(plot = plot_plates_123,
         width = 5, height = 7.5)
```

# Session info for reproducibility

```{r session_info}
sessionInfo()
```



















