---
title: "Plate reader growth curves of GAT201 mutants in RPMI media varying NaHCO3"
author: "Liz Hughes & Edward Wallace"
date: "19/02/2024"
output:
  html_document:
    toc: yes
    toc_depth: 2
---

```{r call_shared_functions_formatting, child = "shared_functions_formatting.Rmd"}
```

# Summary

This experiments tests the effect of sodium bicarbonate (NaHCO3) on growth of Cryptococcus neoformans.

Plate reader growth curves of  WT KN99-alpha and gat201∆ deletion mutant Madhani in RPMI media with no phenol red, and varying NaHCO3 concentrations.

- 3 Bioreps (1 per plate), 3 Techreps each.
- Grow 5ml culture from colony (1 colony = 1 Biorep) in YPD, 30C, 180 rpm ON.
- Seed at OD 600nm = 0.2 , 200 ul per well.
- Run for 1 day / 24 hours at 37C
- With 4 concentrations of NaHCO3: 0mM, 1.5mM, 6mM, 24mM

Data in files, one per biological replicate plate:

- `20240112_SubGoal4_BR1_repeat.xlsx`
- `20240112_SubGoal4_BR2_repeat.xlsx`
- `20240112_SG4_BR3.xlsx`

# Set up data frame of nice strain names, colours, styles.


```{r strain_df}
strain_df <- tibble( Strain      = c("WT", "Gat201 del", "C26", "C30" ),
                     Strain_nice = c("GAT201", "gat201∆m", "GAT201-C1", "GAT201-C2"),
                     colour      = c("grey20", "darkred", "blue3", "cyan4"))

scale_color_strain <- 
  scale_color_manual("Strain",
                     breaks = strain_df$Strain_nice,
                     values = strain_df$colour)
```

```{r media_df}
media_df <- tibble( Media      = c("RPMI-N", "RPMI-1.5mM", "RPMI-6mM", "RPMI-24mM" ),
                    NaHCO3_mM   = c(0, 1.5, 6, 24),
                    NaHCO3_mMfac = as_factor(NaHCO3_mM))
```


# RPMI media analysis, biological replicate 1

## Read in raw data

```{r load_raw_od}
raw_od <- 
  here::here("data", "20240112_SubGoal4_BR1_repeat.xlsx") %>%
  read_platereader_xlsx(start_row = 52)
```

## Read in the Plate map data from csv file RPMI

Also combine with nice strain names for display.

TO DO: nice media info as well

```{r load_platemap}
platemap <- 
  here::here("data", "20240116_SG4_Plate-Setup.csv") %>%
  read_platemap_csv(strain_df = strain_df) 

head(platemap, n=10)

```

## Reshape data and combine with the plate map, pairing them by Well


```{r reshape_annotate}
annotated_od <- 
  reshape_annotate_raw_od(raw_od = raw_od, platemap = platemap)

```


## Plot blank ODs

To find how stable the ODs of the blank wells are.

```{r plot_blank_ODs, fig.width = 5.5, fig.height = 4}
ggplot(data=filter(annotated_od, is.na(Strain)),
       aes(x=Time, y=OD595,  group = Well, colour = Media)) + 
  geom_line() + 
  scale_time_hrs() + 
  scale_y_continuous(limits = c(0, 0.3),
                     expand = c(0, 0),
                     breaks = c(0, 0.05, 0.1, 0.15, 0.2)) +
  labs(title = "Blank wells", x="Time (hours)", y="raw OD595")
```

This shows that...


## Calculate median OD for blank wells


```{r calc_blank_medians}
# filter out columns 1 and 12 as well?
blank_od_summary <- 
  annotated_od %>%
  dplyr::group_by(Media) %>%
  summarise_od_fixed(od.name = "OD595",
                     is.na(Strain)) 

print(blank_od_summary)

blank_od_time <- 
  annotated_od %>%
  dplyr::group_by(Media) %>%
  summarise_od_bytime(od.name = "OD595",
                     is.na(Strain),
                     Well != "H2") 

print(blank_od_time)
```

## Subtract blank OD to make corrected OD RPMI

First check and filter out wells with weird traces...

```{r subtract_blank_OD_RPMI}
bad_blank_wells <- annotated_od %>%
  dplyr::filter(OD595 > 0.18, is.na(Strain)) %>%
  dplyr::pull(Well) %>%
  unique()
  
bad_blank_wells

normalised_od <- 
  annotated_od %>%
  dplyr::filter(!(Well %in% bad_blank_wells)) %>%
  normalise_od_2ways(is.na(Strain))
```

## Plot OD corrected by time-based median summary RPMI

```{r plot_all_stat_summary, fig.width = 5, fig.height = 4}
RPMI_plot <- 
  ggplot_growthcurve_summary(data_od = normalised_od %>%
                               filter(!is.na(Strain)),
                             value_od = "OD_corrected_time",
                             limits_time = c(0,72),
                             limits_od = c(0, 0.1),
                             breaks_od = c(0, 0.05, 0.1)) +
  scale_color_strain(strain_df)

RPMI_plot
```


## Plot OD corrected by time-based median summary RPMI

```{r plot_all_stat_spline_RPMI, fig.width = 5, fig.height = 4}
RPMI_spline_plot <- 
  ggplot_growthcurve_spline(data_od = normalised_od %>%
                              filter(!is.na(Strain)),
                            value_od = "OD_corrected_time",
                            limits_time = c(0,72),
                            limits_od = c(0, 0.1),
                            breaks_od = c(0, 0.05, 0.1)) +
  scale_color_strain(strain_df)

RPMI_spline_plot
```




# Session info for reproducibility

```{r session_info}
sessionInfo()
```



















