---
title: "Plate reader growth curves of GAT201 mutants in RPMI and YPD"
author: "Liz Hughes & Edward Wallace"
date: "02/05/2023"
output:
  html_document:
    toc: yes
    toc_depth: 2
---

```{r call_shared_functions_formatting, child = "shared_functions_formatting.Rmd"}
```

# Summary

Plate reader growth curves of GAT201 mutants

Design:

- Wild-type H99/GAT201, and gat201∆ strains: gat201∆m (Madhani) and gat201∆b (Bahn) in RPMI and YPD at 37 degrees in TECAN plate reader.
- 7 Biological replicates, 3 Technical replicates each. (NB only 2 Techreps for Biorep 7)
- Grow 5ml culture from colony (1 colony = 1 Biorep) in YPD, 30C, 250 rpm ON.
- Seed at OD 600nm = 0.2 , 200 ul per well.
- Run for 3 days/ 72h


# Set up data frame of nice strain names, colours, styles.

```{r strain_df}
strain_df <- tibble( Strain = c("H99", "Gat201(B)", "Gat201(M)", "C26", "C30" ),
                     Strain_nice = c("GAT201", "gat201∆b","gat201∆m", "GAT201-C1", "GAT201-C2"),
                     colour = c("grey20", "magenta3", "darkred", "blue3", "cyan4"))

```


# RPMI media analysis

## Read in transposed data as csv file RPMI

```{r load_raw_od_RPMI}
raw_od_RPMI <- 
  here::here("data", "20200728_EH_PR2_RPMI_WT_Gat201.xlsx") %>%
  read_platereader_xlsx(start_row = 52)
```

## Read in the Plate map data from csv file RPMI

Also combine with nice strain names for display.

```{r load_platemap_RPMI}
platemap_RPMI <- 
  here::here("data", "20200728_EH_PR2_RPMI_Setup.csv") %>%
  read_platemap_csv(strain_df = strain_df) 

head(platemap_RPMI, n=10)

```

## Reshape data and combine with the plate map, pairing them by Well RPMI

```{r reshape_annotate_RPMI}
annotated_od_RPMI <- 
  reshape_annotate_raw_od(raw_od = raw_od_RPMI, platemap = platemap_RPMI)
```

## Plot blank ODs RPMI

To find how stable the ODs of the blank wells are.

```{r plot_blank_ODs_RPMI, fig.width = 5, fig.height = 4}
ggplot(data=filter(annotated_od_RPMI, is.na(Strain)),
       aes(x=Time, y=OD595,  group = Well)) + 
  geom_line() + 
  scale_time_hrs() + 
  scale_y_continuous(limits = c(0, 0.2),
                     expand = c(0, 0),
                     breaks = c(0, 0.05, 0.1, 0.15)) +
  labs(title = "Blank wells RPMI", x="Time (hours)", y="Absorbance at 595 nm")
```

This shows that the OD increases over time, presumably pH change of the phenol red indicator. We correct for this.

## Calculate median OD for blank wells RPMI

Remove well H2 that had a strange reading above.

```{r calc_blank_medians_RPMI}
# filter out columns 1 and 12 as well?
blank_od_summary_RPMI <- 
  summarise_od_fixed(annotated_od_RPMI, od.name = "OD595",
                     is.na(Strain),
                     Well != "H2") 

print(blank_od_summary_RPMI)

blank_od_time_RPMI <- 
  summarise_od_bytime(annotated_od_RPMI, od.name = "OD595",
                     is.na(Strain),
                     Well != "H2") 

print(blank_od_time_RPMI)
```

## Subtract blank OD to make corrected OD RPMI

```{r subtract_blank_OD_RPMI}
normalised_od_RPMI <- 
  normalise_od_2ways(annotated_od_RPMI, is.na(Strain))
```

## Plot OD corrected by time-based median summary RPMI

```{r plot_all_stat_summary_RPMI, fig.width = 5, fig.height = 4}
RPMI_plot <- 
  ggplot_growthcurve_summary(data_od = normalised_od_RPMI %>%
                               filter(!is.na(Strain)),
                             value_od = "OD_corrected_time",
                             limits_time = c(0,72),
                             limits_od = c(0, 0.1),
                             breaks_od = c(0, 0.05, 0.1)) +
  scale_color_strain(strain_df)  +
  labs(title = "RPMI")

RPMI_plot
```


```{r save_png_RPMI}
here::here("results",
           "20200728_EH_PR2_RPMI_WT_Gat201_summaryplot.png") %>%
  ggsave(plot = RPMI_plot,
         width = 5.1, height = 3.75)
```

## Plot OD corrected by time-based median summary RPMI

```{r plot_all_stat_spline_RPMI, fig.width = 5, fig.height = 4}
RPMI_spline_plot <- 
  ggplot_growthcurve_spline(data_od = normalised_od_RPMI %>%
                              filter(!is.na(Strain)),
                            value_od = "OD_corrected_time",
                            limits_time = c(0,72),
                            limits_od = c(0, 0.1),
                            breaks_od = c(0, 0.05, 0.1)) +
  scale_color_strain(strain_df)  +
  labs(title = "RPMI")

RPMI_spline_plot
```


# YPD media analysis

## Read in transposed data as csv file YPD


```{r load_raw_od_YPD}
raw_od_YPD <- 
  here::here("data", "20200807_EH_PR4_YPD_WT_Gat201.xlsx") %>%
  read_platereader_xlsx(start_row = 52)
```

## Read in the Plate map data from csv file YPD

Also combine with nice strain names for display.

```{r load_platemap_YPD}
platemap_YPD <- 
  here::here("data", "20200807_EH_PR4_YPD_Setup.csv") %>%
  read_platemap_csv(strain_df = strain_df) 

head(platemap_YPD, n=10)

```

## Reshape data and combine with the plate map, pairing them by Well YPD

```{r reshape_annotate_YPD}
annotated_od_YPD <- 
  reshape_annotate_raw_od(raw_od = raw_od_YPD, platemap = platemap_YPD)
```

## Plot blank ODs YPD

To find how stable the ODs of the blank wells are.

```{r plot_blank_ODs_YPD, fig.width = 5, fig.height = 3.5}
ggplot(data=filter(annotated_od_YPD, is.na(Strain)),
           aes(x=Time, y=OD595,  group = Well)) + 
  geom_line() + 
  scale_time_hrs() + 
  scale_y_continuous(limits = c(0, 0.2),
                     expand = c(0, 0),
                     breaks = c(0, 0.05, 0.1, 0.15)) +
  labs(title = "Blank wells YPD", x="Time (hours)", y="Absorbance at 595 nm")
```

## Calculate median OD for blank wells for YPD

```{r calc_blank_medians_YPD}
# filter out columns 1 and 12 as well?
blank_od_summary_YPD <- 
  summarise_od_fixed(annotated_od_YPD, od.name = "OD595",
                     is.na(Strain)) 

print(blank_od_summary_YPD)

blank_od_time_YPD <- 
  summarise_od_bytime(annotated_od_YPD, od.name = "OD595",
                     is.na(Strain)) 

print(blank_od_time_YPD)
```

## Subtract blank OD to make corrected OD YPD

```{r subtract_blank_OD_YPD}
normalised_od_YPD <- 
  normalise_od_2ways(annotated_od_YPD, is.na(Strain))
```

## Plot OD corrected by time-based median summary YPD

```{r plot_all_stat_summary_YPD, fig.width = 5, fig.height = 3.5}
YPD_plot  <- 
  ggplot_growthcurve_summary(data_od = normalised_od_YPD %>%
                               filter(!is.na(Strain)),
                             value_od = "OD_corrected_time",
                             limits_time = c(0,72),
                             limits_od = c(0, 0.9),
                             breaks_od = c(0, 0.1, 0.25, 0.5, 0.75)) +
  scale_color_strain(strain_df)  +
  labs(title = "YPD")

YPD_plot
```


```{r save_png_YPD}
here::here("results",
           "20200807_EH_PR4_YPD_WT_Gat201_summaryplot.png") %>%
  ggsave(plot = YPD_plot,
         width = 5.1, height = 3.75)
```

## Plot OD corrected by time-based median summary spline smooth YPD

```{r plot_all_stat_spline_YPD, fig.width = 5, fig.height = 3.5}
YPD_spline_plot <- 
  ggplot_growthcurve_spline(data_od = normalised_od_YPD %>%
                              filter(!is.na(Strain)),
                            value_od = "OD_corrected_time",
                            limits_time = c(0,72),
                            limits_od = c(0, 0.9),
                            breaks_od = c(0, 0.1, 0.25, 0.5, 0.75)) +
  scale_color_strain(strain_df)  +
  labs(title = "YPD")

YPD_spline_plot
```

# Multipanel Figure combining RPMI and YPD results

```{r multifig, fig.height = 4, fig.width = 3.75}
multi_plot <- 
  plot_grid(
    RPMI_spline_plot,
    YPD_spline_plot,
    ncol = 1)

multi_plot
```


```{r save_png}
here::here("results",
           "fig_growthcurves_GAT201_RPMI_YPD.png") %>%
  ggsave(plot = multi_plot,
         width = 3.75, height = 4)
```

```{r save_svg}
here::here("results",
           "fig_growthcurves_GAT201_RPMI_YPD.svg") %>%
  ggsave(plot = multi_plot,
         width = 3.75, height = 4)
```



# Session info for reproducibility

```{r session_info}
sessionInfo()
```



















